---
title: "Reproducible Research: Peer Assessment 1"
author: "Kevin Zembower"
output: 
  html_document:
    keep_md: true
---

```{r setup, echo=FALSE}
# set global chunk options: images will be 7x5 inches
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(digits = 4)
```

## Loading and preprocessing the data

The data on personally-recorded activity was loaded from a
comma-delimited file which was stored on the local drive. This file
was downloaded from the forked project from the instructor, and
decompressed using 'unzip' on the command line. The file was loaded
into R with the commands:

```{r}
df <- read.csv("activity.csv")
```

No further processing was needed to produce the subsequent results.

## What is mean total number of steps taken per day?

The median and mean number of steps taken per day can be produced with these R commands:

```{r}
tapply(df$steps, df$date, mean)
tapply(df$steps, df$date, median)
```

To draw a histogram of the total number of steps taken per day, we can
execute these R commands:

```{r "Barplot_of_Total_Steps_per_Day"}

tot.steps <- tapply(df$steps, df$date, FUN=sum)
barplot(tot.steps)

```

The mean number of steps taken per day is
`r mean(tot.steps, na.rm=TRUE)`
, while the median number of steps taken per day is
`r median(tot.steps, na.rm=TRUE)`.

## What is the average daily activity pattern?

A histogram that illustrates the averae daily activity pattern can be
generated with:

```{r "Total_Steps_per_5-minute_interval"}

int.steps <- tapply(df$steps,
                    df$interval,
                    FUN=mean,
                    na.rm=TRUE
                    )

plot(y=int.steps,
     x=names(int.steps),
     type='l',
     main="Average steps per 5-minute interval",
     xlab='Time at beginning of interval',
     ylab='Total steps',
     )

```

NOTE: This chunk would normally be hidden (not echoed), as it is just
used to calculate a pleasing display of the maximum interval:

```{r}

max.int.steps <- max(int.steps)
hours <- round( as.numeric(names(int.steps[int.steps==max.int.steps])) / 100 )
minutes <- as.numeric(names(int.steps[int.steps==max.int.steps])) %% 100
max.int <- sprintf("%02d:%02d", hours, minutes)

```

The maximum number of steps, `r max(int.steps)`, occurred during the
interval starting at `r max.int`.


## Imputing missing values

### Calculate the total number of missing values

The total number of missing values of 'steps' in the dataset is
`r length(df$steps[is.na(df$steps)])`.

```{r}

       

```

## Are there differences in activity patterns between weekdays and weekends?
