---
title: "Reproducible Research: Peer Assessment 1"
author: "Kevin Zembower"
output: 
  html_document:
    keep_md: true
---

```{r setup, echo=FALSE}
# set global chunk options: images will be 7x5 inches
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(digits = 4)
```

## Loading and preprocessing the data

The data on personally-recorded activity was loaded from a
comma-delimited file which was stored on the local drive. This file
was downloaded from the forked project from the instructor, and
decompressed using 'unzip' on the command line. The file was loaded
into R with the commands:

```{r}

df <- read.csv("activity.csv")

```

No further processing was needed to produce the subsequent results.

## What is mean total number of steps taken per day?

The median and mean number of steps taken per day can be produced with these R commands:

```{r}

tapply(df$steps, df$date, mean)
tapply(df$steps, df$date, median)

```

To draw a histogram of the total number of steps taken per day, we can
execute these R commands:

```{r "Barplot_of_Total_Steps_per_Day"}

tot.steps <- tapply(df$steps, df$date, FUN=sum)
barplot(tot.steps)

```

The mean number of steps taken per day is
`r mean(tot.steps, na.rm=TRUE)`
, while the median number of steps taken per day is
`r median(tot.steps, na.rm=TRUE)`.

## What is the average daily activity pattern?

A histogram that illustrates the averae daily activity pattern can be
generated with:

```{r "Total_Steps_per_5-minute_interval"}

int.steps <- tapply(df$steps,
                    df$interval,
                    FUN=mean,
                    na.rm=TRUE
                    )

plot(y=int.steps,
     x=names(int.steps),
     type='l',
     main="Average steps per 5-minute interval",
     xlab='Time at beginning of interval',
     ylab='Total steps',
     )

```

NOTE: This chunk would normally be hidden (not echoed), as it is just
used to calculate a pleasing display of the maximum interval:

```{r}

max.int.steps <- max(int.steps)
hours <- round( as.numeric(names(int.steps[int.steps==max.int.steps])) / 100 )
minutes <- as.numeric(names(int.steps[int.steps==max.int.steps])) %% 100
max.int <- sprintf("%02d:%02d", hours, minutes)

```

The maximum number of steps, `r max(int.steps)`, occurred during the
interval starting at `r max.int`.


## Imputing missing values

### Calculate the total number of missing values

The total number of missing values of 'steps' in the dataset is
`r length(df$steps[is.na(df$steps)])`.

### Strategy for imputting missing values

My strategy to imput the number of steps is to replace NA's with the
median number of steps for all the days in this interval. I create a
named array of the median of the steps, and replace all NS's with the
median value.

```{r}

steps.median <- tapply(df$steps, df$interval, FUN=median, na.rm=TRUE)

```

### Create an imputted dataset

These R commands will create a new dataset, `df.imput`, with the same
characteristics, but with missing `step` values replace with the
median number of steps for that interval across all days:

```{r imput dataset}

df.imput <- df
df.imput$steps <- ifelse(is.na(df.imput$steps), steps.median, df.imput$steps)

```

### Create a histogram of the imputted values

This R command will plot a histogram of the imputted value of the
total steps taken per day:

```{r "Total_Steps_per_5-minute_interval_imputted"}


tot.steps.imput <- tapply(df.imput$steps, df.imput$date, FUN=sum)
barplot(tot.steps.imput)

```
The mean number of _imputted_ steps taken per day is
`r mean(tot.steps.imput, na.rm=TRUE)`
while the mean number of steps taken per day (_not_ imputted) is
`r mean(tot.steps, na.rm=TRUE)`
.

The median number of _imputted_ steps taken per day is
`r median(tot.steps.imput, na.rm=TRUE)`,
, while the median number of steps taken per day (_not_ imputted) is
`r median(tot.steps, na.rm=TRUE)`.

The impact of imputting data on the mean and median number of steps
taken per day is to _decrease_ both values for the imputted compared to
the non-imputted data.


## Are there differences in activity patterns between weekdays and weekends?
